|

### 2020-05-19 paper.Rmd
Bivariate normals.

>>>>

# Parametrised scaling

Here we describe a parametrised scaling that is controlled by a scaling parameter, rather than fixed given $p$.

We assume that the 2D projected distribution is approximately a bivariate normal with spherical shape (equal variance $\sigma$ on both variables, no co-covariance). In this setting the Mahalanobis distance $r_M$ of any point $(x,y)$ can be written as $r_M = \sqrt{x^2 + y^2}/ \sigma = r / \sigma$, using the polar coordinate $r = \sqrt{x^2 + y^2}$. Using the parametrisation in polar coordinates, we can integrate the distribution to derive the radial CDF $g(r; \sigma) = 1 - e^{-r^2 / 2\sigma^2}$. As before we use $r''=g(r;\sigma)$ and derive the rescaled radius as 
\begin{equation}
r' = f^{-1}(r''; 2, R=1) = \sqrt{1-e^{-r^2 / 2\sigma^2}}.
\end{equation}
Thus we have a rescaling that can be parametrised by $\sigma$, the dependence between $r$ and $r'$ for selected values of $\sigma$ is shown in Figure \ref{fig:radii-normal}.

```{r radii-normal, fig.cap="Relation between $r$ and $r'$ for different values of $\\sigma$. The scaling is approximately linear near the center, but leads to distortion at large radii when $p$ is large.", fig.height=3, fig.width=6, out.width="60%", fig.align = "center"}
# define index as function of c
trans_sig <- function(sig){
  function(r){
    sqrt( 1 - exp(- r^2 / (2* sig^2)))
  }
}

# plot dependence for selected values
ggplot(data = data.frame(x = c(0,1)), mapping = aes(x = x)) +
    stat_function(
        fun = trans_sig(1),
        mapping = aes(color = "ca")) +
  stat_function(
        fun = trans_sig(0.5),
        mapping = aes(color = "cb")) +
  stat_function(
        fun = trans_sig(0.3),
        mapping = aes(color = "cc")) +
    stat_function(
        fun = trans_sig(0.2),
        mapping = aes(color = "cd")) +
      stat_function(
        fun = trans_sig(0.1),
        mapping = aes(color = "ce")) +
  scale_color_manual(name = expression(sigma),
                     values = RColorBrewer::brewer.pal(5, "Dark2"),
                     labels = c("1 (baseline)","0.5", "0.4", "0.2", "0.1")) +
  xlab("r") + ylab("r'") + theme_bw()
